---
title: "Secondary Visualization"
author: "Indigo Edwards"
date: "12/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(lubridate)
library(sf)

cc_full <- read_csv('data/raw/fullCCDataset_2021-12-01.csv')

cc_sites <- read_csv('data/raw/2021-11-18_Site.csv')

us_map <- st_read('data/geographic/states.shp')
  
```

I'm generating a table of the number of usable sites based on a certain number of total surveys across all years within 3 weeks of the summer solstice.

```{r usable surveys, echo = F}

cc_full %>% 
  filter(Year%in% 2018:2021) %>% 
  mutate(
    solstice_jday = if_else(
      Year %in% c(2018,2019),
      true = 172,
      false = 171)) %>% 
  filter(abs(solstice_jday - julianday) <= 21) %>% 
  group_by(SiteFK) %>% 
  summarize(n_surveys = length(unique(ID))) %>%
  summarize(
    '50' = length(which(n_surveys >= 50)),
    '75' = length(which(n_surveys >= 75)),
    '100' = length(which(n_surveys >= 100)),
    '150' = length(which(n_surveys >= 150))) %>% 
  pivot_longer(
    cols = 1:4,
    names_to = 'min_surveys',
    values_to = 'n_sites') %>% 
  knitr::kable(col.names = c('Minimum Total Surveys', 'Number of Sites'))

cc_full %>% 
  filter(Year%in% 2018:2021) %>% 
  mutate(
    solstice_jday = if_else(
      Year %in% c(2018,2019),
      true = 172,
      false = 171)) %>% 
  mutate(
    dfs = abs(solstice_jday - julianday),
    wfs = case_when(
      dfs >= 28 ~ 5,
      dfs %in% 21:27 ~ 4,
      dfs %in% 15:20 ~ 3,
      dfs %in% 8:14 ~ 2,
      dfs < 8 ~ 1)) %>% 
  group_by(SiteFK, wfs) %>% 
  summarize(n_surveys = length(unique(ID))) %>%
  group_by(wfs) %>% 
  summarize(
    '50' = length(which(n_surveys >= 50)),
    '75' = length(which(n_surveys >= 75)),
    '100' = length(which(n_surveys >= 100)),
    '150' = length(which(n_surveys >= 150))) %>% 
  slice(1:4) %>% 
  pivot_longer(
    cols = 2:5,
    names_to = 'min_surveys',
    values_to = 'n_sites') %>% 
  knitr::kable(col.names = c(
    'Weeks from Solstice',
    'Minimum Total Surveys',
    'Number of Sites'))
  
```

For the sake of visualization, we'll use a minimum of 100 surveys.

```{r mapping, echo = F}
cc_100 <- cc_full %>% 
  filter(Year%in% 2018:2021) %>% 
  mutate(
    solstice_jday = if_else(
      Year %in% c(2018,2019),
      true = 172,
      false = 171)) %>% 
  filter(abs(solstice_jday - julianday) <= 21) %>% 
  group_by(SiteFK) %>% 
  summarize(n_surveys = length(unique(ID))) %>%
  filter(n_surveys >= 100) %>% 
  left_join(cc_sites, by = c('SiteFK' = 'ID')) %>% 
  st_as_sf(
    coords = c('Longitude', 'Latitude'),
    crs = st_crs(us_map))

ggplot(us_map) +
  geom_sf() +
  geom_sf(data = cc_100, mapping = aes(color = n_surveys)) +
  theme_void() +
  scale_color_gradient2(
    low = 'purple',
    mid = 'blue',
    high = 'green', 
    midpoint = 1000) +
  labs(color = 'Total Number of Surveys 2018-2021')
  
```

I'm going to visualize where the solstice falls on abundance curves over each summer.

```{r abundance curves, echo = F, message = F}

rect_df <- tibble(
  Year = rep(2018:2021,3),
  weeks = factor(c(
    rep(2,4),
    rep(3,4),
    rep(4,4))),
  xmin = c(
    158,158,157,157,
    151,151,150,150,
    143,143,142,142),
  xmax = c(
    186,186,185,185,
    193,193,192,192,
    200,200,199,199),
  ymin = rep(0,12),
  ymax = rep(Inf,12))

cc_full %>% 
  filter(Year%in% 2018:2021) %>% 
  mutate(
    solstice_jday = if_else(
      Year %in% c(2018,2019),
      true = 172,
      false = 171)) %>% 
  mutate(arths = if_else(is.na(arthID),0,1)) %>% 
  group_by(SiteFK, Year, julianday, solstice_jday) %>% 
  summarize(abundance = sum(arths)) %>% 
  ggplot() +
  geom_line(aes(x = julianday, y = abundance, group = factor(SiteFK))) +
  facet_wrap(~Year, scales = 'free') +
  geom_vline(aes(xintercept = solstice_jday), color = 'red') +
  geom_rect(
    data = rect_df,
    mapping = aes(
      xmin = xmin,
      xmax = xmax,
      ymin = ymin,
      ymax = ymax,
      fill = weeks),
    alpha = .4) +
  labs(
    color = 'Number of Weeks from Solstice',
    x = 'Julian Day',
    y = 'Number of Arthropods Observed',
    title = 'Abundance Curves with Solstice-Centered Sampling Periods')

```

Next steps:

1. Correct table of values to be all surveys within x or fewer weeks of solstice.
